SymPy

\begin{verbatim}
%load_ext sympyprinting
\end{verbatim}

\begin{verbatim}
from __future__ import division
from sympy import *
\end{verbatim}

We now show a few examples of SymPy's capabilities to show its usefulness as a
general computer algebra system. In SymPy symbols have to be created
explicitely. Here we create the symbols ($x$, $y$, $z$):

\begin{verbatim}
x, y, z = symbols("x y z")
\end{verbatim}

Once these symbols have been created, symbolic expressions can be created and
manipulated. Here is an example involving a rational number, $\pi$, $i$, $e$ and
basic arithmetic operations:

\begin{verbatim}
Rational(3,2)*pi+exp(I*x) / (x**2+y)
\end{verbatim}
$$\frac{3}{2} \pi + \frac{e^{\mathbf{\imath} x}}{x^{2} + y}$$

Expressions can be evaluated numerically using SymPy's arbitrary precision
arithmetic algorithms:

\begin{verbatim}
exp(pi*sqrt(163)).evalf(35)
\end{verbatim}
$$262537412640768743.99999999999925007$$

Here is an example of expanding and refactoring a polynomial:

\begin{verbatim}
((x+y)**2*(x+1)).expand()
\end{verbatim}
$$x^{3} + 2 x^{2} y + x^{2} + x y^{2} + 2 x y + y^{2}$$

\begin{verbatim}
factor(_)
\end{verbatim}
$$\left(x + 1\right) \left(x + y\right)^{2}$$

Derivatives can be taken of expressions:

\begin{verbatim}
diff(cos(x**2)**2/(1+x), x)
\end{verbatim}
$$- 4 \frac{x \operatorname{sin}\left(x^{2}\right) \operatorname{cos}\left(x^{2}\right)}{x + 1} - \frac{\operatorname{cos}^{2}\left(x^{2}\right)}{\left(x + 1\right)^{2}}$$

We emphasize that these examples are merely representative of its vast
capabilities. We encourage the reader to consult the SymPy documentation for a
more complete set of examples.

Python is an object oriented language and SymPy leverages that to build reusable
objects for symbolic mathematics. Internally, SymPy creates instances of these
objects and stores them in an expression tree. In this example, we construct a
simple expression and then display its expression tree and SymPy objects using
the \verb:srepr: function:

\begin{verbatim}
e = x + 2*y
srepr(e)
\end{verbatim}

\begin{verbatim}
Add(Symbol(&apos;x&apos;), Mul(Integer(2), Symbol(&apos;y&apos;)))
\end{verbatim}

This combination of object orientation and expression trees makes SymPy easily
extensible; once new objects are created they can immediately be used in
arbitrary expressions.

